Local to global N input modelling

### 2025-06-06 manuscript.qmd
Nicer map on Robinson/Mercator

@@

worldmap <- ne_countries(scale = "medium", returnclass = "sf")

p1 <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

@@

```{r patchwork map}
#| fig-cap: |
#|   Possibly use a patchwork map to show spatial results?
#| fig-height: 9
#| fig-width: 9.5

worldmap <- ne_countries(scale = "large", returnclass = "sf")

AUS_bounds <- c(144, 149.5, -39.75, -44)
CAN1_bounds <- c(-132, -122, 47.75, 54.25)
CAN2_bounds <- c(-70, -54, 43, 48.5)
CHI_bounds <- c(-77.5, -62.5, -56, -25)
EUR_bounds <- c(-26, 30, 51, 71)

p_ALL <- ggplot(data = worldmap, aes(geometry = geometry)) +
  geom_sf(fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(colour = mean_temp), size = 1.5) +
  coord_sf() +
  theme_classic()

@@

p_Aus <- p_ALL + coord_sf(xlim = AUS_bounds[1:2], ylim = AUS_bounds[3:4], expand = FALSE) + no_margins()
p_Eur <- p_ALL + coord_sf(xlim = EUR_bounds[1:2], ylim = EUR_bounds[3:4], expand = FALSE) + no_margins()
p_Chi <- p_ALL + coord_sf(xlim = CHI_bounds[1:2], ylim = CHI_bounds[3:4], expand = FALSE) + no_margins()
p_Can <- p_ALL + coord_sf(xlim = CAN1_bounds[1:2], ylim = CAN1_bounds[3:4], expand = FALSE) + no_margins()
p_USA <- p_ALL + coord_sf(xlim = CAN2_bounds[1:2], ylim = CAN2_bounds[3:4], expand = FALSE) + no_margins()

subplot <- (p_Eur/p_Aus) + plot_layout(heights = c(1.5, 1))
finalplot <- (p_Can + p_USA) / (p_Chi + subplot) + 
  plot_layout(widths = c(1, 1.5), heights = c(1, 2)) &
  theme(plot.margin = margin(0, 0, 0, 0))
finalplot
```

>>>>

```{r custom fuctions}
#| code-summary: Function to create sf polygons from bounding boxes

create_boxes <- function(box_list, label_positions, offset_deg) {
  boxes <- list()
  labels <- list()

  for(i in seq_along(box_list)) {
    name <- names(box_list)[i]
    coords <- box_list[[i]]
    pos <- label_positions[i]

    # Create polygon from bounding box coordinates
    box_coords <- matrix(c(
      coords[1], coords[3],  # lonmin, latmin
      coords[2], coords[3],  # lonmax, latmin  
      coords[2], coords[4],  # lonmax, latmax
      coords[1], coords[4],  # lonmin, latmax
      coords[1], coords[3]   # close polygon
    ), ncol = 2, byrow = TRUE)

    # Create sf polygon
    poly <- st_polygon(list(box_coords))
    boxes[[i]] <- st_sfc(poly, crs = 4326)

    # Calculate label position based on specification
    label_coords <- get_label_position(coords, pos, offset_deg[i])
    label_point <- st_point(c(label_coords[1], label_coords[2]))
    labels[[i]] <- st_sfc(label_point, crs = 4326)
  }

  # Combine into sf objects
  boxes_sf <- st_sf(
    name = names(box_list),
    letter = LETTERS[1:length(box_list)],
    geometry = do.call(c, boxes)
  )

  labels_sf <- st_sf(
    name = names(box_list), 
    letter = LETTERS[1:length(box_list)],
    geometry = do.call(c, labels)
  )

  return(list(boxes = boxes_sf, labels = labels_sf))
}

# Helper function to calculate label positions
get_label_position <- function(coords, position, offset) {
  lonmin <- coords[1]
  lonmax <- coords[2]
  latmin <- coords[3]
  latmax <- coords[4]

  # Parse position string
  parts <- strsplit(position, "_")[[1]]
  vertical <- parts[1]    # "top" or "bottom"
  horizontal <- parts[2]  # "left" or "right"
  location <- parts[3]    # "inside" or "outside"

  # Base coordinates for corners
  if(vertical == "bottom") {base_lat <- latmin} else {base_lat <- latmax}
  if(horizontal == "right") {base_lon <- lonmax} else {base_lon <- lonmin}

  # Apply offset for outside positioning
  if(location == "outside") {
    if(horizontal == "left") {base_lon <- base_lon-offset} else {base_lon <- base_lon+offset}
    if(vertical == "top") {base_lat <- base_lat + offset} else {base_lat <- base_lat - offset}
  }

  return(c(base_lon, base_lat))
}

lim_robin <- function(lons = c(144.0, 149.5), lats = c(-39.75, -44.00)) {
  coords <- data.frame(lon = lons, lat = lats) %>%
    st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = "+proj=robin") %>% 
    st_coordinates()
  list(
    xlims = range(coords[, "X"]),
    ylims = range(coords[, "Y"])
  )
}

```

@@

worldmap <- ne_countries(scale = "medium", returnclass = "sf") %>% st_transform(crs = "+proj=robin")
graticules <- st_graticule(worldmap, lon = seq(-180, 180, 30), lat = seq(-90, 90, 30))

inset_boxes <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 47.75, 54.25),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-26, 30, 51, 71),
  CHI = c(-77.5, -62.5, -56, -25),
  AUS = c(144, 149.5, -39.75, -44)
)
labels_spec <- c(CAN1 = "bottom_left_outside", CAN2 = "bottom_right_outside", EUR = "bottom_left_outside", CHI = "top_left_outside", AUS = "bottom_left_outside")
labels_offset <- c(CAN1 = 1, CAN2 = 2, EUR = 1.75, CHI = 1.75, AUS = 3.5)

box_data <- create_boxes(box_list = inset_boxes, # see function create_boxes above
                         label_positions = unname(labels_spec),
                         offset_deg = unname(labels_offset))
boxes_robinson <- st_transform(box_data$boxes, crs = "+proj=robin")
labels_robinson <- st_transform(box_data$labels, crs = "+proj=robin")
farm_temp_robinson <- st_sf(farm_temp_means) %>% st_transform(crs = "+proj=robin")

p1 <- ggplot() +
  geom_sf(data = graticules, color = "gray80", size = 0.3) +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_robinson, aes(color = mean_temp), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "horizontal", position = "bottom",
    label.position = "bottom", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 25,barheight = 1.5
  )) +
  geom_sf(data = boxes_robinson, fill = NA, color = "darkred", size = 1.65) +
  geom_sf_text(data = labels_robinson, aes(label = letter), 
               color = "darkred", size = 5, fontface = "bold", 
               hjust = 0.5, vjust = 0.5) +
  theme_void()

@@

```{r patchwork-map-mercator}
#| fig-cap: |
#|   Possibly use a patchwork map to show spatial results?
#| fig-height: 9
#| fig-width: 9.5

@@

inset_boxes_sm <- list(   # lonmin, lonmax, latmin, latmax
  CAN1 = c(-132, -122, 48.5, 54),
  CAN2 = c(-70, -54, 43, 48.5),
  EUR = c(-23, 28.5, 52, 71),
  CHI = c(-78, -62, -55.5, -27),
  AUS = c(144, 149, -40.75, -43.75)
)

worldmap <- ne_countries(scale = "large", returnclass = "sf")
p1 <- ggplot() +
  geom_sf(data = worldmap, fill = "white", color = "dimgray") +
  geom_sf(data = farm_temp_means, aes(color = mean_temp, geometry = geometry), size = 1) +
  coord_sf() +
  scale_color_viridis_c(option = "turbo") +
  guides(col = guide_colourbar(
    title = "Mean temperature (°C)",
    direction = "vertical", position = "right",
    label.position = "right", title.position = "top", 
    title.vjust = 1, title.hjust = 0.5,
    frame.colour = "black", ticks.colour = "black",ticks.linewidth = 2,
    barwidth = 1.5, barheight = 20
  )) +
  prettyplot() +
  labs(y = "Latitude", x = "Longitude")

p_Aus <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["AUS"]][1:2], ylim = inset_boxes_sm[["AUS"]][3:4])  +
  draw_label("E", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["AUS"]][1], y = inset_boxes_sm[["AUS"]][3], 
             hjust = 0, vjust = 1)
p_Eur <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["EUR"]][1:2], ylim = inset_boxes_sm[["EUR"]][3:4]) +
  draw_label("C", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["EUR"]][1], y = inset_boxes_sm[["EUR"]][4], 
             hjust = -1, vjust = 1)
p_Chi <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CHI"]][1:2], ylim = inset_boxes_sm[["CHI"]][3:4]) +
  theme(legend.position = "right") +
  draw_label("D", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CHI"]][1], y = inset_boxes_sm[["CHI"]][4], 
             hjust = 0, vjust = 0)
p_Can1 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN1"]][1:2], ylim = inset_boxes_sm[["CAN1"]][3:4]) +
  draw_label("A", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN1"]][1], y = inset_boxes_sm[["CAN1"]][4], 
             hjust = -1, vjust = 0)
p_Can2 <- p1 + 
  coord_sf(xlim = inset_boxes_sm[["CAN2"]][1:2], ylim = inset_boxes_sm[["CAN2"]][3:4]) +
  draw_label("B", size = 14, fontface = "bold", 
             x = inset_boxes_sm[["CAN2"]][1], y = inset_boxes_sm[["CAN2"]][4], 
             hjust = 0, vjust = 1)

grid_1 <- cowplot::plot_grid(p_Can1, p_Can2, ncol = 2, rel_widths =  c(1, 1.35))
grid_2 <- cowplot::plot_grid(p_Eur,  p_Aus,  nrow = 2, rel_heights = c(1, 1))
grid_3 <- cowplot::plot_grid(grid_2, p_Chi,  ncol = 2, rel_widths =  c(1, 1))

cowplot::plot_grid(
  grid_1, grid_3,
  nrow = 2,
  rel_heights = c(1, 1.5)        # relative row heights
)
```
